### 2016-11-28 chapter1.tex
Corrections made to ISR Section

Because convection maps require data covering as wide a range of magnetic local time sectors as possible, SuperDARN radars are continuously operational, usually in a common mode.  This creates a vast database of high-quality measurements of small-scale FAIs, which is very useful for other studies on irregularity occurrence and plasma structuring.

\subsection{Incoherent Scatter Radar}
\label{sec:isr}
Similar to CSRs, incoherent scatter radars (ISRs) are ground based radars that are used to probe the ionosphere.  However, they use a fundamentally different technique to observe structuring, which allows them to measure different things than CSRs.  Where as CSRs receive backscatter from large, "coherent" density structures in the ionosphere, ISRs operate by radar beams scattering due to the random thermal motion of electrons in the ionosphere.  The radar receivers then detect a spectra of frequencies with different power.  Analysis of this spectra will reveal the plasma frequency, and hence, the electron density, electron and ion temperature, and LoS velocity \citep{Gordon1958}.

The first ISR was built in Arecibo, Puerto Rico in 1958.  Since then, at least 8 other ISRs have been deployed around the world at equitoral, mid-, and high-latitudes.  The most recent advancement has been the development of Advanced Modular Incoherent Scatter Radars (AMISR), of which there are currently three in use, one at the Poker Flat Rocket Range, just north of Fairbanks, AK and two in Resolute Bay, Canada.  The advantage of these new systems is that they are electronical steerable, unlike older ISRs which consisted of a large dish that had to be moved manually to change the beam direction.  This allows beams to be transmitted in many different directions at a very high time cadence, giving the radar the capability to make measurements in multiple directions almost simultaniously.  This is very important for creating 2D maps of ionospheric conditions and tracking how they change in time.

>>>>

Because convection maps require data covering as wide a range of magnetic local time sectors as possible, SuperDARN radars are continuously operational, usually in a common mode.  This creates a vast database of high-quality measurements of small-scale FAIs, which is very useful for other studies on irregularity occurrence and plasma structuring.  In the studies presented here, data was primarily used from the SuperDARN radars located at Rankin Inlet, Canada (RKN) and McMurdo Station, Antarctica (MCM).  The have been shaded in Figure \ref{fig:superdarnmap}.

\subsection{Incoherent Scatter Radar}
\label{sec:isr}
Similar to CSRs, incoherent scatter radars (ISRs) are ground based radars that are used to probe the ionosphere.  However, they use a fundamentally different technique to observe structuring, which allows them to measure parameters of the actual ionospheric plasma.  Where as CSRs receive backscatter from relatively large, "coherent" density structures in the ionosphere, ISRs operate by radar beams scattering due to the random thermal motion of electrons in the plasma.  The radar receivers then detect a spectra of frequencies with different power \citep{Gordon1958}.

The scattering cross section for a volume of N electrons in the ionosphere is given by \(\sigma_n = N \sigma_e\) for radio waves with a wavelength small compared to the Debye length of the plasma, where \(\sigma_e\) is the scattering cross section of a single electron \citep{Gordon1958,Fejer1960}.  The relationship to the plasma Debye length is important because at scale sizes smaller than the Debye length, the plasma is not capable of organized motion, so the motion of the elections is only due to their own thermal energy and not larger scale plasma waves.  The signal returned from this kind of scatter was postulated to have a very low power and a very broad spectral width, however after the first ISR was built, the spectrum was found to be much narrower than previously expected \citep{Evans1969}.  This can be attributed to the effect that the much slower ions have on free electrons in the plasma through ion acoustic waves \citep{Bowles1958}.  The practical results of this is that ISRs can use radio waves of a much longer wavelength than originally anticipated, on the order of \(\sim\)1 m.

Analysis of the returned power spectrum reveals the electron density, electron and ion temperatures, and LoS ion velocity \citep{Evans1969,Rishbeth1985,Nicolls2007}.  Using a model of chemical composition at different altitudes, the average ion mass can also be found.  In addition, 3D convection velocities and electric field vectors can be derived from LoS ion velocity measurements in some of the more modern ISRs by assuming both \(\vec{E}\) and \(\vec{V}_E\) are constant across the radar's FoV \citep{Heinselman2008}.  Each instrument must be calibrated with known electron density measurements to account for noise and system constants \citep{Nicolls2007}.  This is done using the plasma frequency measured from either the plasma line of the ISR spectra during summer daytime period or ionosonde measurements for all other times \citep{Bachivan2010,Themens2014}.

The first ISR was built in Arecibo, Puerto Rico in the 1950s \citep{Gordon1958}.  Since then, at least 8 other ISRs have been deployed around the world at equitoral, mid-, and high-latitudes.  The most recent advancement has been the development of Advanced Modular Incoherent Scatter Radars (AMISR), of which there are currently three in use, one at the Poker Flat Rocket Range, just north of Fairbanks, AK and two in Resolute Bay, Canada.  The advantage of these new systems is that they are electronical steerable, unlike older ISRs which consisted of a large dish that had to be moved manually to change the beam direction.  This allows beams to be transmitted in many different directions at a very high time cadence, giving the radar the capability to make measurements in multiple directions almost simultaneously \citep{Nicolls2007a,Nicolls2007b,Bachivan2010}.  This is very important for creating 2D maps of ionospheric conditions and tracking how they change in time \citep{Semeter2009,Dahlgren2012a,Dahlgren2012b}.  AMISR systems typically have two types of pulses, a long pulse (LP) with 72 km range gates for F-region studies and an alternating code pulse (AC) with 4.5 km range gates gates for E region studies.  The exact number and configuration of beams depends on the radar mode used and can be changed easily due to the system's electronic steering.  In the common WorldDay mode, there are typically 11 beams and data is collected at \(\sim\)1 min intervals, although this can change slightly between different renditions of the WorldDay mode.

One of the most important advantages of ISR systems over CSRs is the ability to directly measure electron density in the ionosphere.  However, because ISRs are much scarcer and are only typically run on in WorldDay mode a few days out of every month, data is not as widely available.  In addition, ISRs do not actually measure small-scale coherent plasma structures like the HF SuperDARN network does.  However, they can directly measure large-scale structures and image density variations on scales larger that 100 km (limited by the number of beams and range gate size).  For these region, it is often useful use a combination of CSR and ISR measurements in ionospheric plasma structuring studies, when possible.  The primary ISR used in the following studies is the north face of the AMISR system at Resolute Bay, Canada, RISR-N.  The FoV of RISR-N overlaps that of the RKN SuperDARN radar, making it ideal for these types of comparison studies.
